---
title: "Intro to time series wrangling and viz"
author: "Laurel Abowd"
date: "1/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(tsibble)
library(feasts)
library(slider)
```

## Read in data

```{r}
toolik <- read_csv(here("data", "toolikweather.csv"))
```
## Convert this to a tsibble so that R understands it is time series data

```{r}
toolik_ts <- toolik %>% 
  mutate(date = mdy(date)) %>% 
  as_tsibble(key = NULL, index = date)
```

```{r}
ggplot(data = toolik_ts, aes(x = date, y = mean_airtemp)) +
  geom_line()
```

# Use the "index_by()" function to aggregate time series info by intervals we specify 

```{r}
# Calculate the mean monthly air temperature
toolik_month <- toolik_ts %>% 
  index_by(yr_mo = ~yearmonth(.)) %>% 
  summarize(monthly_mean_temp = mean(mean_airtemp, na.rm = TRUE))

ggplot(data = toolik_month, aes(x = yr_mo, y = monthly_mean_temp)) +
  geom_line() +
  facet_wrap(~month(yr_mo, label = TRUE))

# Can see strong seasonality in the data set!

```

# Aggregate by week 
```{r}
toolik_weekly <- toolik_ts %>% 
  index_by(weekly = ~yearweek(.)) %>% 
  summarize(weekly_airtemp = mean(mean_airtemp, na.rm = TRUE))
head(toolik_weekly)
```

## Filtering time series 

Use 'filter_index()' to specify range of time series to keep or exclude. 

```{r}
toolik_ts %>% 
  filter_index("2000-06" ~ "2001-10")
```

```{r}
toolik_ts %>% 
  filter_index("2006-04-10" ~ "2006-05-15")
```

## Seasonplots

```{r}
toolik_ts %>% 
  filter(year(date) > 2014) %>% 
  gg_season(y = mean_airtemp)

toolik_month %>% 
  gg_season(y = monthly_mean_temp)
```

## Seasonal subseries plot

```{r}
toolik_month %>% 
  gg_subseries(monthly_mean_temp)
```

## Find the moving average (use slider package in R)

```{r}
set.seed(2021)
test <- rnorm(n = 100, mean = 40, sd = 10)


slide(test, ~.x, .before = 2, .after = 2)

```


```{r}
roll_toolik_15 <- toolik_ts %>% 
  mutate(ma_15d = as.numeric(slide(toolik_ts$mean_airtemp,
                                   mean, 
                                   .before = 7,
                                   .after = 7)))
# Use as.numeric b/c slide returns values as list.

ggplot() +
  geom_line(data = toolik_ts, aes(x = date, y = mean_airtemp), size = 0.2, color = "gray") +
  geom_line(data = roll_toolik_15, aes(x = date, y = ma_15d)) +
  theme_minimal()
# Plotted original data and then the moving average data on top
```

## Create the ACF 

```{r}
toolik_ts %>% 
  ACF(mean_airtemp) %>% 
  autoplot()

toolik_month %>% 
  ACF(monthly_mean_temp) %>% 
  autoplot()
```

## Decomposition by STL

```{r}
toolik_dec <- toolik_month %>% 
  model(STL(monthly_mean_temp ~ season(window = Inf)))

components(toolik_dec) %>% autoplot()

```

